---
title: "Lab 3: "
author: "POLS 3316, Fall 2023, University of Houston, Instructor: Tom Hanna"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

## Part 1: Scalars, Vectors, Matrices and Arrays

### What is a matrix?

A *matrix* is a rectangular array or table of numbers, symbols, or expressions, arranged in rows and columns, used to represent some mathematical object or property. 

Practical examples of *matrices*? Spreadsheets and tables

#### Matrixes in R

In R, every *dataframe* or *table* or similar *object* is a matrix.

For an example of building a matrix in R, we need to look at scalars and vectors. 


### What is a scalar?

A scalar is a single number. A scalar is matrix with dimension 1x1, 1 column by 1 row. A scalar is also a vector with only 1 element.

#### Scalars in R

To create a scalar in R, we can either assign a null value and add the value later, or we can assign a value when we creat it:

```{r}
#create an ojbect called a and assign the null value
#making it a null value scalar
a <- c()
cat('a is a null scalar ')
a

#create an object called b and assign the value 5, making 
#b a scalar object
b <- 5
cat('b is a scalar with value 5')
b

#assign the value 1 to a, making it a scalar with value 1
a <- 1
cat('Now I have given a the value 1')
a
```



### What is a vector?

A vector is a list of numbers. It is either a single column or a single row. It is a matrix with dimension 1 x n . 

#### Vectors in R?

To create a vector in R, we can assign a series of values either by just listing them or through a variety of functions including operations on other vectors:

```{r}

#create object c and assign <- a series of numbers 
#using c() creating a vector object
C <- c(1,2,4,5,6,7,8,9,10)
cat('C is a vector ')
C

#create a vector D and assign the value of multiplying 
#vector C by scalar b
D <- C*b
cat('D is a vector created by multiplying b * C ')
D

```

```{r}

#Create a vector E and assign the value
#for the seq function - in this case
#sequence starting at -5 and going to 5
#by steps of 0.5
E <- seq(-5, 5, by = .5)
cat('E is a sequence of numbers from -5 to 5, by 0.5s ')
E

```

```{r}

#create the vector object F and 
#assign a function of E and the 
#normal distribution (more on this later)
F <- dnorm(E, mean = 2.5, sd = 0.5)
cat('F is a function of E and the normal distribution ')
F


```

## Things to know

- Multiplying anything times a scalar is simply multiplying every element by that scalar. It's almost always feasible. 
- To do an operation on two vectors they must be the same length
- Multiplying matrices is a complex version of multipying vectors. - If you see errors related to matrix/vector size, it usually means you are multiplying two matrixes who aren't a match in column and row size. 

If you need more help with matrix operations in the future (not this class), you have two options:

https://www.mathsisfun.com/algebra/matrix-multiplying.html

or 

**ask for help from someone with a good grasp of basic matrix algebra**

## Back to matrices in R

We built a scalar object and a vector object. How do we build a matrix object?

Let's use two of the vectors we built, C and D:

```{r}
#create an object called data_cd and assign <- 
#vector C and D as a matrix in R's data frame format
#using the data.frame() function
data_cd <- data.frame(C,D)

#Show the first six lines of data_cd
head(data_cd)

```




## Part 2: Measures of dispersion in R          

We're going back to the arrest data. We can use the R functions *var()* to get the **sample** *variance* and *sd()* to get the **sample** *standard deviation*. The numbers we figured were the **population** numbers. The difference in the formulas is that for the *variance*, the population version divides by the number of observations *n* and the sample version divides by *n-1*. We'll discuss the theoretical difference in detail this week, but for now the difference is this: In most cases we are only working with a sample of data and we are using it to estimate the full "population" or "universe" of data. The slight difference in formula is to account for this.  

For now, let's start by looking at the Arrest Data *sample variance* and *sample standard deviation.*


To get **sample variance** use the R function **var().**


```{r}

#loading arrests_data built into R
#create an object called arrests_data
#assign the 2 and 3 columns [2:3] from the 
#USArrests dataset 
arrests_data <- USArrests[2:3]

#create an object called urban_pop_variance
#and assign the value created by applying the 
#var() function to the UrbanPop variable from the 
#arrests_data dataframe
urban_pop_variance <- var(arrests_data$UrbanPop)
cat('Urban Pop Percent Variance')
#print urban_pop_variance to the screen
urban_pop_variance

#create an object called urban_pop_variance
#and assign the value created by applying the 
#var() function to the Assault variable from the 
#arrests_data dataframe
assaults_variance <- var(arrests_data$Assault)
cat('Assaults Variance')

#print assaults_variance to the screen
assaults_variance

#print the mean of Assault variable from
#arrests_data to the screen
mean(arrests_data$Assault)

```



To get **sample standard deviation** either take the **square root of sample variance** or use the R function **sd()**.

```{r}
#the cat('') command will print to the screen
cat('Standard deviations')
#create object urban_pop_sd1 and assign the 
#value of sd(arrests_data$UrbanPop)
#sd() is the R function for sample standard deviation
urban_pop_sd1 <- sd(arrests_data$UrbanPop)
urban_pop_sd1

#Do the same thing by taking the square root of the variance
#instead of using sd()
urban_pop_sd2 <- sqrt(var(arrests_data$UrbanPop))
urban_pop_sd2

```


To get **population variance** multiply the R function var() times (n-1)/n:  var(x) * (n-1)/n

```{r}

#Converting sample variance to population variance
#Create object urban_pop_popvar and assign the 
#math operation multiplying the sample variance times
#(n - 1)/n where n = 50
urban_pop_popvar <- var(arrests_data$UrbanPop) * (50-1)/50
cat('Urban Pop population variance:')

#print the result
urban_pop_popvar

#same for Assault variable
assaults_popvar <- var(arrests_data$Assault) *(50-1)/50
cat('Assaults population variance:')
assaults_popvar

```


To get **population standard deviation**, the only option is to compute the population variance and takes its square root: **sqrt(populationvariance)** *You can't get it by multiplying the R function at least easily.*

```{r}

#take the square root of the population variance
#to get the population standard deviation
sqrt(urban_pop_popvar)
sqrt(assaults_popvar)

```





