---
title: "Matrices and Lists in R"
date: today
author: "POLS3316, Instructor: Tom Hanna, Fall 2023, University of Houston"
format: 
        revealjs:
                self-contained: true
                transition: convex
                theme: [moon, custom.css]
                logo: logo.png
                footer: "GOVT2306, Fall 2023, Instructor: Tom Hanna"
---


## Matrices and Lists

- The core of this lesson is matrices and lists
- We will use matrices

                - We will use R matrices occasionally
                - We will use objects based on matrices almost constantly
                - data frames are mathematically matrices but different from R matrices

- We will not use lists much if at all
- Why am I covering them?

                - Lists are important core objects in R
                - They are not that complicated
                - Understanding them has other value
                - Lists underly a lot of code from packages


## Scalars and Vectors

The building blocks of matrices: scalars and vectors. 

- Scalars
                - A single unit of information
                - A number
                - A letter
                - A symbol
                - Zero dimensional, geometrically a point
                
Vectors

                - A series of scalars
                - Of the same data type
                - A series of numbers - 5,7,21,13,11,1
                - A series of letters - z,f,a,l,w,n
                - 1 dimensional, geometrically a line with length but no width
                
                
## Creating a scalar in R

- create the object name
- assign 
- a single value

```{r echo=TRUE}
#create a scalar named my_scalar and assign the value 75
my_scalar <- 75

my_scalar #print the scalar to the screen or output

```

## Creating a scalar in R

- It's fairly conventional to name scalars with one or two letters instead of a long name
- The reason will shortly be apparent

```{r echo=TRUE}
# Create a scalar named X1 with value 43
 
X1 <- 43

X1 # print the value to the screen


```




## creating a vector in R

Ways to do this:

- import some data that has one column
- enter a series of values in parentheses using c()
- enter a series of scalars using c()
- enter a sequence created with an R function

We will skip the first method

## Creating a vector in R: series of values

- Create the object name
- assign with the assignment operator
- list the values in parentheses separated by commas

```{r echo=TRUE}
# create an object named my_vector and assign a series of numbers
# 
my_vector <- c(15,23,12,18,73)

my_vector #print my_vector to screen

```
## Creating a vector in R: series of scalars

First we need to create some scalars to use

```{r echo=TRUE}
X2 <- 123
X3 <- 17
X4 <- 99
X5 <- 42
X6 <- 31

X2
X3
X4
X5
X6

```



                
## Matrices

                - A series of vectors
                - 
                
- 



## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this 

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
