---
title: "VDem data subsetting"
author: "Tom Hanna"
format: 
        html:
                self-contained: true
---

## Loading VDem data

The first two commented lines (lines that start with #) are for installing the VDem data. To use that data on your own computer, delete the leading # and run the lines. Do not delete the second hashtag in the middle of the line. The material beyond that is still a comment. 

```{r echo=TRUE}

#install.packages("devtools") #uncomment and run one time
#devtools::install_github("vdeminstitute/vdemdata") #uncomment and run one time

library(vdemdata) #load vdem library

vdem_data <- vdem #load vdem data into an object called vdem_data for local use


```

## search topics: Civil society

This is an example of the code to find variables on a given keyword

```{r echo=TRUE}

# The next line creates an object called civ_soc_vars and uses a 
# command from the VDem package to find all VDem variabes related to
# Civil Society and assign them to the object as a data frame.
# 
civ_soc_vars <- find_var("Civil Society") 

# This returns columns 4 through 9 of the dataf rame to the screen 
# when the file is rendered.
# 
civ_soc_vars[(4:9)]


```

## search topics: conflict

```{r}

#The next two lines do the same thing for variables with
#keyword conflict

conflict_vars <- find_var("conflict")
conflict_vars[(4:9)]

```


## Search topics: income

```{r}

gdp_vars <- find_var("gdp")
gdp_vars[(4:9)]

```



## Search topics: development

```{r}

development_vars  <- find_var("development")
development_vars[(4:9)]

```
## Search topics: immigration

```{r}

immigration_vars <- find_var("immigration")
immigration_vars[(4:9)]

```




## Search topics: democracy

```{r}

democracy_vars <- find_var("democracy")
democracy_vars[(4:9)]


```

## Search topics: property rights

```{r}

property_vars <- find_var("property")
property_vars[(4:9)]


```



## Search topics: repression

```{r}

repression_vars <- find_var("repression")
repression_vars[(4:9)]


```

## Search topics: democracy

```{r}

democracy_vars <- find_var("democracy")
democracy_vars[(4:9)]


```


## Search topics: ideology

```{r}

ideology_vars <- find_var("ideology")
ideology_vars[(4:9)]


```


## Search topics: opposition

```{r}

opposition_vars <- find_var("opposition")
opposition_vars[(4:9)]

```