---
title: 'Lecture 15: August 2, 2022'
output:
  html_document:
    df_print: paged
---

## Confidence Intervals

+ Review $X^2$ and t-test 
+ General hypothesis testing: scores and critical values
+ Confidence intervals 

## $X^2$ and t-test

+ Note on formulas: Numbers in numerator or denominator of a fraction all get solved together as if in parentheses. PEMDAS is shorthand, but the P should really be G for "grouping symbols" and includes:  parentheses ( ), brackets [ ], braces \{ \}, and even fraction bars.

+ Terms

                + null hypothesis
                + alternative hypothesis
                + critical value (cutoff point)
                + critical value is found using degrees of freedom and desired probability or alpha. (In z-scores, only probability is needed.)
                + probability or p - alpha level (new term) - confidence level (new term)
                

+ test scores and critical values - if the test score is greater than the critical value, reject the null

                + The critical value tells us the minimum value
                + Above that value, the probability that the null hypothesis is true is equal to or less than the target so we reject
                
 $$\\[5in]$$      
 
 ## Test scores
 
 Monday Quiz and other questions
 
 
 $$\\[5in]$$      
 
 ## Confidence intervals
 
 ### What is a confidence interval?
 
+ It's the range where we could expect to find the true measurement based on the probability level we specify if we had the full population data instead of just the sample. 
+ It's the range where we expect to find the population parameter given the sample statistic
+ If we say the 95% confidence interval of the mean is from 45 to 55, it means we are 95% certain that the true population mean lies somewhere in that range. 
+ The confidence intervals are based on the thing we actually measured, the probability level, and the distribution. 
+ If we know the population parameters (especially standard deviation) we can create simple CI using 68-95-99 rule or accurate CI using z-scores
+ For small samples or uncertain population standard deviation we use t-scores to adjust for both these uncertainties 

              
 $$\\[5in]$$     
 
## Using t-scores to find confidence intervals

 
+ For small samples or uncertain population standard deviation we use t-scores to adjust for these uncertainties 
 
                + If we wanted to get a 95% confidence interval with 68-95-99, we would look for a value 2 standard deviations either side of the statistic we want to find a CI for 
                + With t-test scores we divide the standard deviation by the square root of sample size and multiply times a t-table critical value instead.
                + We can get the critical value using the qt() function in R as in the examples
                
+ $CI_\mu$ = $\bar{x} \pm t \frac{s}{\sqrt{n}}$

+ $CI_\mu$ Confidence interval of the population mean
+ $\bar{x}$ Sample mean
+ s = sample standard deviation
+ n = sample size


 $$\\[5in]$$        
 
 ## What about the normal distribution?

The 68-95-99 rule gives an oversimplified CI. 

For normal distributions, the actual formula is:

$CI_\mu = \bar{x} \pm z \frac{\sigma}{\sqrt{n}}$

$CI_\mu$ Confidence interval of the population mean
$\bar{x}$ Sample mean
$sigma$ population standard deviation
$n$ is the sample size

+ We can get the z-score from a z-table or from the qnorm() function in R


 $$\\[5in]$$    
 


 
## Examples:

We'll go through a couple of example computations, then I'll show you some examples as we actually use them in lab.

Sample mean: 50, sample standard deviation 5, sample size 18. Looking for a 95% confidence interval. We need the t-score:

```{r}

qt(.025,17, lower.tail = FALSE)

```

Then we can fill in the formula:

+ $CI_\mu$ = $\bar{x} \pm t \frac{s}{\sqrt{n}}$

```{r}

```




Using the z-score:

Sample mean: 50
Population standard deviation: 5
Sample size: 500

We need a z-score for 95% (.025% above and .025% below):

```{r}

qnorm(.025, lower.tail = FALSE)

```
Then we apply the formula:

$CI_\mu = \bar{x} \pm z \frac{\sigma}{\sqrt{n}}$


```{r}

```


$$\\[5in]$$                    
                
## Is this just about means?

NO!

The biggest use I see of confidence intervals is confidence intervals for **regression coefficients**

                $y = \alpha + \beta X + \epsilon$
                
If we find that the $\beta$ in the equation defining the relationship between our explanatory (x) and dependent (y) variables is 0.9, the confidence interval can tell us that it's actually 95% likely that the true relationship is between 0.88 and 0.92. 

Even better, we can graph it, which we will see in a few minutes.


$$\\[5in]$$     


 https://www.statisticshowto.com/probability-and-statistics/confidence-interval/#WhatisCI
 
 https://www.statisticshowto.com/probability-and-statistics/statistics-definitions/confidence-level/
 
 https://www.statisticshowto.com/95-percent-confidence-interval/
