---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
Lab 13 is regression examples. Some will be n this notebook and some will be in other files.

```{r}
library(here) #I have to use this to get data from #other folders easily

#load data from student_outcomes.csv in data #directory using here() function and read.csv() #function
#assign to data frame object student_outcomes
student_outcomes <- read.csv(here("data","student_outcomes.csv"))


#look at first six rows with head() and variable #details with #summary()

head(student_outcomes)

summary(student_outcomes)


```
There are three variables.

Campname is a character type and is a categorical variable. It is actually just the observation names.

Outcomes.dropdown is a continuous numeric variable with range 0 to 1 and seems to be a good candidate for a Y/dependent/outcome variable.

Total.PPE is a continuous numeric variable with range 6022 to 12997. I don't know what it means, but realistically that is our only option for an X variable. 



```{r}

#create a model named outcomes_model and assign the 
#lm() function with Outcomes.Dropdown as the Y and #Total.PPE as the X. data = student_outcomes to #reference the correct dataframe

outcomes_model <- lm(Outcomes.Dropdown ~ Total.PPE, data = student_outcomes)

#look at just the coefficients
outcomes_model$coefficients
```

```{r}
#look at a summary of the model

summary(outcomes_model)


```
```{r}

#Use stargazer to make a nicer table for #presentation
library(stargazer)
stargazer(outcomes_model, type = "text")
```
Generate a prediction with 95% confidence interval for the mean

```{r}
#create a data frame with the desired value
new_data <- data.frame(Total.PPE <- mean(student_outcomes$Total.PPE))


predict(outcomes_model,new_data, interval = "confidence")

```


```{r}

#plot this with ggplot
library(ggplot2)

#create scatterplot with fitted regression line
ggplot(student_outcomes, aes(Total.PPE, Outcomes.Dropdown)) + 
  geom_point() +
  stat_smooth(method = "lm")



```

Show predictions for multiple points (the summary data points - min, 1st quartile, median, mean, 3rd quartile, max)

```{r}
#create a data frame with the desired value
Total.PPE <- c(6022,8625,9448,9571,10413,12997)

new_data2 <- data.frame(Total.PPE)


predict(outcomes_model,new_data2, interval = "confidence")

```