---
title: 'July 15 Quiz: POLS 3316'
output:
  html_document:
    df_print: paged
---

## Quiz 2 and Problem Set 1 combined

### Due 11:59 PM July 16

## Part 1 - Complete before leaving class Friday

**Run code in R Studio or R Studio Cloud. Download the finished HTML or PDF file and upload to Blackboard by the due date for full credit**

**The areas with ``` followed by {r} and ending with ``` are code chunks. The Run dropdown arrow in the upper right of this box give you options for running the code. Run *Current Chunk* runs just the piece you are working on. After you have finished, you can *Run all* then save the result. Where you see Preview (above just left of center) there is a dropdown arrow. If you select "Knit to HTML" or "Knit to PDF", you can produce a copy that can be downloaded and submitted in Blackboard.**

**IF all else fails, screenshot your work and submit that in Blackbaord.**

For more help with this RMarkdown notebook here are some resources:

Cheatsheet - https://rmarkdown.rstudio.com/lesson-15.HTML

Code Chunks - https://rmarkdown.rstudio.com/lesson-3.html


1. a. (bonus) Create a vector named *my_vector* containing the sequence (1,2,3,4,5). Find the sum using the vector. 


```{r}
#use the assignment operator <- to assign the value
#use the c binding method - c() - to create the sequence
my_vector <- c(1,2,3,4,5)


```

1. Write the code to create an object named *object_sum* whose value is the sum of the sequence of numbers (1,2,3,4,5) from problem 1 and then return the value of that object to the console. (*Hint: you'll use the assignment operator to assign an operation on the right side to object_sum on the left side. Hint2: You can type in a math operation or you can use your code from question 1. Hint 3: To return the value, I've already written the code, all you have to do is uncomment it.*)

```{r}

#create it with a math operation
object_sum <- 1 + 2 + 3 + 4 + 5

#second option

#create it using the sum() function on my_vector object
object_sum2 <- sum(my_vector)

```

2. What is the sum of the sequence of numbers (1,2,3,4,5)? (You can code in R for practice by pressing CTRL-ALT-I to create a code "chunk" in the answer spot, or you can add the answer and type it.)

Answer: 15

3. How many observations (n) are in the sequence (1,2,3,4,5)?

Answer: 5 **You can either count them, or look next to my_vector where is shows [1:5] indicating that it is a sequence of 5 values**

3. a. Write simple code to create an object called *object_n* and assign this number to it. (*Hint: if you use more than the object name, the assignment operator, and a single digit number you are working too hard!*)

```{r}

#This createst the object
object_n <- 5

#This prints the object to the screen to verify it

object_n

```

4. Write the code to create a new object named *object_mean* whose value is the mean of the above values using *object_sum* and the number of observations *object_n*? (Hint: You'll use the assignment operator with the *object_mean* on the left side and a math operation using the other objects on the right side.)

```{r}

#this creates the object
object_mean <- object_sum / object_n

#this returns the object as output to the screen (prints, returns, #echoes are various ways of saying this)
object_mean

```


5. Get help with the *mean()* function in R in the next code chunk. If you need help for problem 6, you can type the same thing in the Console (bottom left): 

```{r}

#To get help with a function, you can use ?
#For some non-base functions, you may need to use ??

?mean



```

6. Now write the code to find the mean of the above values using the *mean()* function in R? (*Hint: If you did the bonus question, 1a it will save you time here*)


```{r}

#To just print the mean to the screen

mean(my_vector)


```

7. Write the code to find the sample variance of the above observations using the objects you created (*object_n, object_mean, object_sum*) and the sequence of values? (*Hint: Creating an object with the sample variance will help with question 8 but remember to return the object's value if you do.*)

**In the answer, I do this two ways and use the R function to check, if it's right all three values should agree**


```{r}


#code to do this
my_sample_variance <- sum((my_vector - object_mean)^2)/(object_n -1)
my_sample_variance

#alternate method

msv2 <- ((1 - object_mean)^2 + (2 - object_mean)^2 + (3 - object_mean)^2 + (4 - object_mean)^2 + (5 - object_mean)^2)/(object_n - 1)
msv2

var(my_vector)

```

8. Write code to find the standard deviation? 

**There are two simple ways to do this:**

```{r}

my_sd <- sqrt(my_sample_variance)
my_sd

my_sd2 <- sd(my_vector)
my_sd2

```

9. What is the median of sequence of values from question 1. 

**with 5 numbers, just counting to the middle number is easy and you can do that to check. It's also an acceptable answer. Here is the R code:**

```{r}

median(my_vector)

```

10. What is the mode of  0, 1, 1, 2, 3, 5, 8, 13, 21, 34. 

Answer: 1 **All that you need to do is count. This will be on the midterm and final and should be an easy point!**

11. What is your favorite food?  

Answer: any answer

12. What is your Github username?  

Answer: tomhanna-uh is mine

13. What is the name of your project data set? 

Answer: If you haven't picked, do so soon!

14. What is the source of your data set?  

Answer: varies - Kaggle, Dataverse, government agencies are good sources. Built in R Datasets are acceptable.

15. What would you like to learn from the data set in one or two sentences? (How to use data/R studio/statistitcs is an acceptable answer, but if you want to know more I can help you.)

Answer: Varies. Start thinking about that compelling answer for the first paragraph of the project - why is this data interesting? What can we learn from it?

16. (Optional) How many variables does your data have? How many observations does your data have? 

Answer: Varies by dataset. 

17. (Optional) Does your data have observations across multiple time periods?

Answer: Varies by dataset. 

### Part 2

*For the remaining problems, you can use math operations or R functions. You can use the methods above (creating objects, etc.) or you can just run the functions on the raw numbers. *

18. What is the sum of the sequence (5,10,15,20,25,30,35,40)

```{r}

vector_2 <- c(5,10,15,20,25,30,35,40)
sum_vector_2 <- sum(vector_2)
sum_vector_2

```

19. How many observations (n) are in the sequence in question 18?

Answer: 8

20. What is the mean of the sequence in question 18?

```{r}

n_vector_2 <- 8
mean_vector_2 <- sum_vector_2 / n_vector_2
mean_vector_2

```

21. What is the sample variance of the sequence in question 18?

```{r}

#create an object with the variance
vector_2_variance <- sum((vector_2 - mean_vector_2)^2)/(n_vector_2 - 1)
#return the object to the screen
vector_2_variance

#check with the R function
var(vector_2)

```

22. What is the sample standard deviation of the sequence in question 18?

```{r}

#use the square root of the sample variance
vector_2_sd <- sqrt(vector_2_variance)
vector_2_sd

#check with the R functino
sd(vector_2)
```

23. What is the median of the sequence in question 18?

**Since it is an even number, it is halfway between the two middle numbers - 20 and 25. You can either do the math (20 + 25)/2 to find the midpoint or use the R function. If this happens on a test, you'll need to do the math.**

```{r}
median(vector_2)
```


24. What is the difference between a sample variance and a population variance? (One or two sentences.)

The population variance is the number we get when we have data for every occurrence. The sample variance is what we get when we only have data for a subset of the population, a sample of the population. 

25. What is the sum of the sequence (3,7,5,5,6,7,7,9)

```{r}
#create a vector of the sequence
my_seq <- c(3,7,5,5,6,7,7,9)
#use the sum function on the vector to create a new vector with #the answer
sum_my_seq <- sum(my_seq)
#return the new vector with the answer
sum_my_seq

#Option 2
3 + 7 + 5 +5 + 6 + 7 + 7 + 9


```

26. What is the mode of the sequence? 7 

27. How many observations (n) are in the sequence? 8

28. What is the mean of the sequence?

```{r}

49/8

mean(c(3,7,5,5,6,7,7,9))

mean(my_seq)

```
