---
title: |
    | Hypothesis Testing in R
    | POLS 3316, Fall 2023
author: "Tom Hanna"
format:
        html:
                self-contained: true
                code-fold: true
                code-summary: "Show the code"
                theme: vapor
---


## Chi square

#### Checking scores

Use pchisq - first argument is the score, second argument is degrees of freedom

Our critical value is .05. For any value less than .05, we reject the null hypothesis.

So, to get the p-value for X^2 score of 4, df of 1:

```{r}

1 - pchisq(4, df = 1)

# or

pchisq(4, df = 1, lower.tail = FALSE)

```
For this value we:  Reject the null hypothesis that the two variables are independent.


```{r}

1 - pchisq(1, df =  3)

# or 

pchisq(1, df = 3, lower.tail = FALSE)


```
For this value we: Fail to reject the null hypothesis that the two variables are independent.

## Running chisquare on data



```{r}

#Reading in some sample data
#Author DataFlair
data_frame <- read.csv("https://goo.gl/j6lRXD")  #Reading CSV
table(data_frame$treatment, data_frame$improvement)



```

Format is:

chisq.test(first_variable,second_variable, correct = FALSE) if all cells have at least 5 observations

chisq.test(first_variable,second_variable, correct = TRUE) if any cell has less than 5 observations

```{r}

chisq.test(data_frame$treatment,data_frame$improvement, correct = FALSE)
```




## paired sample t-test

Second Example Data:


Group 1: (12.2, 14.6, 13.4, 11.2, 12.7, 10.4, 15.8, 13.9, 9.5, 14.2)

Group 2: (13.5, 15.2, 13.6, 12.8, 13.7, 11.3, 16.5, 13.4, 8.7, 14.6)

Are the means of the two groups different?

This is a two-tailed test because there is no direction.

If the question was, is Group 1 greater than Group 2, then it would be a one-tailed test.

## Loading the data as vectors
                
```{r}
#Group 1 as vector named Group 1 using the c() function
Group1 <- c(12.2, 14.6, 13.4, 11.2, 12.7, 10.4, 15.8, 13.9, 9.5, 14.2)

# Group 2 as vector named Group 2 using the c() function
Group2 <- c(13.5, 15.2, 13.6, 12.8, 13.7, 11.3, 16.5, 13.4, 8.7, 14.6)


```
 
#combining the vectors into a data frame
 
```{r}

# combining the vectors into a dataframe using the data.frame() function
data <- data.frame(Group1,Group2)

#look at the data using the head() function
head(data)
```
            
```{r}
# summary of the data using the summary() function
# gives a summary of every variable in the dataframe
summary(data)
```

## Two ways to solve

### Option 1: Work it out step by step.

1. Create a variable with the sample differences.
          
```{r}

# ceate a third vector named difference in the dataframe
data$difference <- data$Group1 - data$Group2

# Summary of the difference variable
summary(data$difference)

```
         
2. Get the number of observations (n), mean of the sample differences and standard deviation of the sample differences and assign to objects:

```{r}

# mean of the difference variable
n <- as.numeric(length(data$Group1))

mean_diff <- mean(data$difference)
mean_diff

sd_diff <- sd(data$difference)
sd_diff
```

3. Fill in the formula using the objects created above


$t = \frac{\bar{x}_{diff}}{\sigma_{diff}/\sqrt{n}}$

```{r}

mean_diff/(sd_diff/sqrt(n))


```

4. Check the t-score against the table.


The function is 

2 * pt(q, df, lower.tail=TRUE) for a two-tailed test

or

pt(q, df, lower.tail=TRUE) for a one-tailed test

where:

q: The t-score

df: The degrees of freedom



```{r}

2 * pt(-2.27202, 9, lower.tail=TRUE)

```

More on reading t-tables plus 1- and 2- tailed tables here:

[https://www.statisticshowto.com/tables/t-distribution-table/](https://www.statisticshowto.com/tables/t-distribution-table/)

### Option 2:  using R functions

Or we can use the R function

```{r}

#Vectors go in x,y
#this is a paired test, so paired = TRUE
test1 <- t.test(Group1,Group2, paired = TRUE)
test1

```