---
title: 'Lab 6: July 20'
author: 'POLS 3316, Summer 2022, University of Houston, Instructor: Tom Hanna'
output:
  html_document:
    df_print: paged
---

##Probabilities with R frequency distribution functions

In this lab, we use the distribution functions from R for calculating probabilities.

1. Create a variable x with values (-2,-1,0,1,2)

```{r}

x <- c(-2,-1,0,1,2)

```

2. What is the probability of getting a value less than a given value (to the left of that value in the distribution) in a normal distribution with mean 0 and standard deviation 1? 

*mean = 0, standard deviation = 1 is the default for the normal distribution functions in R*

First, let's look graphically at what that means:


```{r, figures-side, fig.show="hold", out.width="50%", echo=FALSE}
#This code is to generate the data for the two plots
rand.norma<- rnorm(100000)

# calculate CDF 
CDF <- ecdf(rand.norma)

#This code draws the two plots, first the Red Histogram then the #Cumulative Distribution plot
#the xaxp=c(-4,4,16)  argument sets the x-axis tick marks to run from -4 to 4 and to have a total of 16 tick marks 

hist(rand.norma, breaks = 200, freq = TRUE, main = "Normal Distribution \n sd = 1, 100,000 random draws", xaxp=c(-4,4,16) , xlab = 'x', col = "red")

plot(CDF, main = "Cumulative Distribution Function \n normal dist, sd = 1, 100,000 random draws")

```

**What would we expect with each of those values?**

$$\\[4in]$$
## Example with minus -2: 

```{r}
#pnorm gives the probability to the left of (x)
pnorm(-2)

```
3. Create a vector that contains the probabilities for all values of x and return the vector.

```{r}

pnorm_x <- pnorm(x)
pnorm_x

```

4. Now create a variable “y” with values: (0,1,2,5,8,10,15,20)

```{r}

y <- c(0,1,2,5,8,10,15,20)

```


5. Using a binomial distribution, calculate the value of the cumulative distribution function at a given value of y. Example: 0

```{r}

pbinom(0,size=20,prob=.2)

```

6. Now create a vector with the binomial probabilities of all y values. 

```{r}

pbin_y <- pbinom(y,size=20,prob=.2)

```

7. Repeat Question 6 but using the Poisson and Uniform Distributions - commands are *ppois()* and *puni* - create vectors named vector_y_pois and vector_y_uni

```{r}



```

## Calculating probabilities with the functions: 

8. Probability that a normal random variable with mean 22 and variance 25 lies between 16.2 and 27.5. 

First, to make this clearer, I'm going to plot this distribution function by simulating it with 100000 randomly generated numbers:



```{r}
#First, we need the standard deviation
sqrt(25)

#we use the pnorm function - the probability distribution function #for the normal distribution
#The first argument is the vector (or scalar) we want to compute
#The second argument is the mean of the distribution (22)
#The third argument is the standard deviation 
# First, I am going to compute the values at each point and assign to vectors
point_1 <- pnorm(27.5,22,sd=5)
point_2 <- pnorm(16.2,22,sd=5)

#We subtract one result from the other
answer8 <- point_1 - point_2
answer8

#It is possible to just subtract one function from the other
answer_8a <- pnorm(27.5,22,sd=5) - pnorm(16.2,22,sd=5)
answer_8a

```
9. Probability that a normal random variable with mean 22 and variance 25  is greater than 29 

```{r}
#We use the same function for a value of 29
#and subtract from 1

1 - pnorm(29,22,sd=5)

```

10. If we wanted to know the probability for **less than** 17 instead of greater than:

```{r}

pnorm(17,22,sd=5)

```

11. And if we wanted to know the probability that it was less than 15 or greater than 25 (mutually exclusive possibilities):


```{r}
#The parentheses around the two parts aren't strictly necessary
#I just wanted to separate the two probabilities
(pnorm(15,22,sd=5)) +  (1-pnorm(25,22,sd=5))

```




