---
title: 'Lab 8: July 22'
author: 'POLS 3316, Summer 2022, University of Houston, Instructor: Tom Hanna'
output: html_notebook
---

Quiz and Problem Set Notes:

First: getting updated copies of Labs and Lectures in R Studio Cloud, then some general notes, then the quiz part of yesterday's lab. 

object - something that is created and named that R can perform operations on. Once we create it, we can ask R to perform operations on it just by using the name. Examples are vectors, dataframes, and functions. Exception: If the Environment is cleared or you start a new session. In this case, you may need to rerun the code stored in the script, notebook, or markdown file. 
 
 Environment (sometimes Global Environment) - this is the place where objects you create are stored. You can see a list of the Environment contents in the right quarter of the screen in R Studio. 
 
 Types of objects: vectors, scalars, matrices including dataframes, functions, models, and others
 
 Loading data: the command to load a CSV file is read.csv() . I've been using it with a function called "here" in order to load data from a separate data directory. If the data is in your main project directory, you don't need here(). All you'll need is something like this:
 
 my_data <- read.csv("the_actual_data_file_name.csv") 
 
 Once you run that, my_data is now an object that you can call by name.

 
 Now, if you want to pull a particular variable out of that dataframe, you have to know its name, too.
 
 For example, if I want to call the variable my_variable from my_data I use the $ like this
 
 my_data$my_variable
 


## Part I: Loops

Loops are funtions, or sets of instructions, which tell the computer to repeat an operation or set of operations several times. 
There are two kinds of loops in R, *for loops* and *while loops*. There is also a family of functions, the apply functions that act similary to loops with added advantages - *apply(), lapply(), tapply(), sapply(), vapply(),* and *mapply()*. Starting with understanding basic loops helps with understanding how the computer processes repeated instructions. It is also possible to nest loops within loops, which can be very powerful. Today we are going to cover basic simple, unnested *for loops* and *while loops*

#### For loops

A very simple loop tells R to start at some number, print it, then move to the next in the range:

```{r}
#i is an index that R uses to track where it is in the loop
#i increases by 1 on each pass until it reaches the end of the range
#1:5 is the range. The loop starts at i = 1 and runs until i = 5
#All this loop does is print the value of i
for (i in 1:5) {
  print(i)
}

```

You can also use a loop to perform vector operations. For example:

```{r}
#first, we need a vector to operate on

my_vector <- c(1,2,3,4,5)

#and a second vector to get the output
my_vector2 <- c()

for (i in 1:5) {
  my_vector2 <- c(my_vector2,my_vector * i)
}
print(my_vector2)

```

And we can loop through elements in a dataframe:

```{r}
#first we need some data
library(here) # load the here library
#create ojbect gotdata and have read.csv load the 
##file GOT_screentimes.csv in data subdirectory
gotdata <- read.csv(here("data","GOT_screentimes.csv")) 

for (name in gotdata) {
        Character <- (gotdata$name)
        Actor <- (gotdata$portrayed_by_name)
}
new_df <- data.frame(Actor,Character)
played_by <- " was played_by "

for (i in 1:191) {
      nth <- paste0(new_df$Character[i],played_by,new_df$Actor[i]) 
      print(nth)
}

```

###While loops

```{r}
#set the start point
i <- 1

#tell R to run the loop as long as is less than 29
while (i < 29) {
        #tell it what to do
        nth <- paste0(new_df$Character[i],played_by,
                      new_df$Actor[i]) 
      print(nth)
      #tell it to add 1 to i (otherwise, i never gets bigger and the loop runs forever)
      i = i + 1
}
```


## Part II: Frequency Distributions (Cont) and Correlations

