---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

## paired sample t-test

Second Example Data:


Group 1: (12.2, 14.6, 13.4, 11.2, 12.7, 10.4, 15.8, 13.9, 9.5, 14.2)
Group 2: (13.5, 15.2, 13.6, 12.8, 13.7, 11.3, 16.5, 13.4, 8.7, 14.6)

## Loading the data as vectors
                
```{r}
Group1 <- c(12.2, 14.6, 13.4, 11.2, 12.7, 10.4, 15.8, 13.9, 9.5, 14.2)
Group2 <- c(13.5, 15.2, 13.6, 12.8, 13.7, 11.3, 16.5, 13.4, 8.7, 14.6)


```
 
#combining the vectors into a data frame
 
```{r}

data <- data.frame(Group1,Group2)

head(data)
```
            
```{r}
summary(data)
```

Work it out step by step.

First, create a variable with the differences.
          
```{r}

data$difference <- data$Group1 - data$Group2

summary(data$difference)

```
         
Get the mean and standard deviation of the differences and assign to objects:

```{r}
n <- length(data$Group1)

mean_diff <- mean(data$difference)
mean_diff

sd_diff <- sd(data$difference)
sd_diff
```

Fill in the formula:

```{r}

mean_diff/(sd_diff/sqrt(n))


```

To find the critical value using R
qt(p/2, df, lower.tail=FALSE)

where:

p: The significance level to use (/2 since it is a two tailed test)
df: The degrees of freedom
lower.tail: If TRUE, the probability to the left of p in the t distribution is returned. If FALSE, the probability to the right is returned. Default is TRUE.

```{r}

qt(.05/2, 9, lower.tail=FALSE)

```

More on reading t-tables plus 1- and 2- tailed tables here:

[https://www.statisticshowto.com/tables/t-distribution-table/](https://www.statisticshowto.com/tables/t-distribution-table/)

Or we can use the R function

```{r}

#Vectors go in x,y
#this is a paired test, so paired = TRUE
test1 <- t.test(Group1,Group2, paired = TRUE)
test1

```

## Simulations in R

### Why do simulations in R?

+ The real world is random (stochastic)
+ But we know data often falls within regular distributions (normal, binomial, Poisson, uniform, etc.)
+ R can generate random samples drawn from known distributions
to help us model the real world

### Where do we start?

First, for most simulations, we use the *set.seed()* function in R to choose our random seed:

```{r}
#Set seed assures that our random outcome will be the same if we repeat the trial
set.seed(2)

```

This actually takes the randomness out to an extent - it means we pick a specific random outcome that will be the same as long as we set the same seed. 

Next, we determine what we want to simulate.

We can simulate 25 random observations from a uniform distribution with a range from 0 to 10 as follows:

```{r}

#runif is a random number from a uniform distribution
#The first argument (25) is the number of observations we want to
#generate. Min and max set minimum and maximum values for the #range to be generated
runif(25,min=0,max=10)

```

We can also generate round numbers:

```{r}
#using the tidyverse - dplyr
#
library(dplyr)

#round with digits = 0 gives integers
runif(25,min=0,max=10) %>% 
  round(.,digits = 0)

#or using base r
#round with digits = 2 yields 2 decimal places
round(runif(25,min=0,max=10), digits = 2)

```

So, far we've just created bunches of numbers on the screen. We actually need them assigned as objects. Here, we'll create vectors of equal length using random numbers from several distributions:

```{r}
#25 random numbers from 0 to 10, with 2 digits, drawn from a uniform distribution
vector_a <- round(runif(25,min=0,max=10), digits = 2)

#25 random numbers drawn from a normal distribution, mean =5 , #standard deviation = 2
vector_b <- rnorm(25, mean = 5, sd = 2)

#25 random numbers drawn from a poisson distribution with a lambda
#(mean and variance) equal to 5
vector_c <- rpois(25, lambda = 5)

```

We can also use the sample function:

```{r}

#We are selecting randomly from a sequence (seq) from 1 to 10
#25 times, with replacement
vector_d <- sample(seq(1,10), 25, replace = TRUE)


```

We can use sample to draw randomly from a vector or dataframe, which can be really useful in simulating with real data and we can mix it up by using unequal probabilities for the draws:

```{r}

sample(vector_b, 25, replace = TRUE, prob =c(0.1,0.25,0.4,0.25,0.1,0.25,0.4,0.25,0.1,0.1,0.25,0.4,0.25,0.4,0.25,0.1,0.25,0.4,0.25,0.1,0.1,0.25,0.4,0.25,0.25))
```

With a dataframe, we can use the row and column indexes to randomly select. For example, we can randomly select 10 of the 50 states from the USArrests dataset:

```{r}

# step 1: create an index vector for the elements/rows 
index <- seq_len(nrow(USArrests))
 
# step 2: sample the index vector
set.seed(12)
#to obtain a random sample of 10
sample_index <- sample(index,10)
# step 3: to show the sampled elements/rows 
USArrests[sample_index,]

```

A common example is to simulate die rolls. Here we use a simulation to check the probability of rolling an 8:

```{r}
#setting a seed 
set.seed(75)

#creating a vector with the possible die rolls
die <- 1:6

# sample 10000 times with replacements for dice 1
roll1 <- sample(die,10000,replace = TRUE,prob=NULL)


# sample 10000 times with replacements for dice 2
roll2 <-  sample(die,10000,replace=TRUE,prob = NULL)


# the combined value of roll1 and roll2
outcomes = roll1 + roll2

# the probability of getting a 8 (actually the proportion of the outcomes that are 8)
mean(outcomes == 8)

```

## Aside from random number generation

Aside from just randomly generating numbers and randomly drawing variables, we can use the other R functions including loops (and the *apply* functions), conditional statements (*if* and *ifelse*) to simulate different scenarios with data. 