---
title: "Lab 2: "
author: "POLS 3316, Fall 2023, University of Houston, Instructor: Tom Hanna"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

## Math functions in R

In lecture I used some math functions. The first one I used to add a sequence of numbers:

```{r}
#create an object called sum_7 using the math operation
sum_7 = 5 + 6 + 7 + 8 + 9 + 10

#print sum_7 to the screen
sum_7 

```

The "sum_7 <-" was something I did to create an object I could use and reuse. You can see that object now in the upper right corner. The "sum_7" by itself was necessary to get the answer to return (or print or echo) to the screen. We'll come back to this in more detail. But you can also just type in the math function if you just need an answer that you don't want to reuse:

```{r}

#just perform the operation without creating an object
5 + 6 + 7 + 8 + 9 + 10

```

### Basic math functions

- Addition +
- Subtraction -
- Multiplication *
- Division /

We'll try a few on here and then you can try a few on your own.

How would I:

- Add 4 and 5?

- Subtract 10 minus 5?

- Divide 10 by 2?

- Multiply 10 by 2?

### Basic plus math functions

- Exponents ^
- Square root of x - sqrt(x)
- Modulus (the remainder from division) %%
- Integer division %/%

Examples:

- 4 to 3rd power (or cubed)
- 2 squared (or second power)
- Find the remainder of 5 divided by 2
- Divide 5 by 2, rounding down
- Square root of 16

```{r}

#4 to the power of 3
4^3

#2 to the power of 2 (2 squared)
2^2

#5 modulus 2
5%%2

#5 integral division 2
5 %/% 2

#square root of 16
sqrt(16)

```

The modulus is probably not something you have seen before. It's useful in programs for things like checking if a number is even or odd. If you divide by 2 and it's even, the modulus will be zero. If it's odd, the modulus will be one.:

```{r}

#5 modulus 2
5%%2

#4 modulus 2
4%%2

10%%2

248682%%2

7456835%%2

```

R follows the standard order of operations including parentheses:

```{r}
#two math operations demonstrating use of parentheses in R
5 + 6/2

(5+6)/2
```


Try a few problems. You can just type in the console area (lower left):

- 5 plus 5
- 5 to the 5th power
- 10 + 10
- (6 + 6)/2
- 6 + 6/2
- 10 minus 20

### Relational operators

- Less than <
- Greater than >
- equal to == (Note: a single equal sign does something different)
- Less than or equal to <=
- Greater than or equal to >=
- Not equal to !=

These are usually used in programs or functions, some of which may be very short. When you use them, R simply tells you if the statement, which is called a *condition* in R is TRUE or FALSE

```{r}

#Testing relational operators
#True statements will return "True"
#False statements will return "False"
5 < 3
5 > 3
5 != 3
5 == 3
#5 = 3

```

```{r}
#More relational operators
3 > 3
3 >= 3
3 <= 3
3 != 3

```

You can find more information on these here:

[R Operators: Arithmetic, Relational, Logical, Operational and more](https://www.datamentor.io/r-programming/operator/)

Some other useful math functions are:

- sum(x) - returns the sum of a vector of numbers 
- log(x) - gives the natural log of x
- exp(x) - gives the exponential of x (inverse of the natural log)
- round(x, n) - rounds *x* to the nearest *n* decimal places
- signif(x, n) - gives *x* to the nearest *n* significant figures


Try a few on your computer by typing in the console:

- Round 10.12345 to the nearest 2 decimal places
- Find the natural log of 12
- Find the exponential of 2.484907 -
- Is the natural log of 10 equal to 2.5






**The answers**

```{r}

#Round 10.12345 to the 2nd decimal place
round(10.12345,2)

#natural log of 12
log(12)

#exponential function e of 2.484907
exp(2.484907)

#relational operator ==
#R will return true or false for the question
#Does the log(10) == 2.5 ?
log(10) == 2.5

```

## Descriptive Stats using Math functions

### Assignment operator

- The assignment operator in R is <-
- It tells R that the *object* on the left side has the value on the right side
- This allows you to do things with that value later

Example, create a new object and give it the value 7:

```{r}

#The left side creates a new object called my_new_object
# <- is the assignment operator that tells R to make the
# left side equal to the right side
my_new_object <- 7

```
That's great, but what can I do with it? 

First, I can return (or print or echo) the value to the screen:

```{r}
#this is all you need to do to have R print an object to the
#screen - type the object name
my_new_object
```
I can create an even newer object and assign a value to it based on the first object and some math operation. Then I can return the result to the screen:

```{r}

#create even_newer_object and make it equal to
#my_new_object plus 10
even_newer_object <- my_new_object + 10

#print even_newer_object to the screen
even_newer_object

```
Apply this to finding the mean of:

5,6,7,12,13,14

The mean is just the sum of the numbers divided by the number of observations. First, I'll create an object called *my_sum* and assign the value of the math operation to get the sum. I'm also going to return that number to the screen, just as a check to make sure it looks right:

```{r}

#create object my_sum with value equal to sum of the digits
my_sum <- 5 + 6 + 7 + 12 + 13 +14

#return my_sum to screen
my_sum

```

To get the mean, we divide that by the total observations. **We can just count those - 6 **

First, I'll just do this as a simple math problem, using the answer from the previous part. 

```{r}

#just figuring the mean (average) by hand from 
#the sum divided by the number of values in the series
57 / 6

```


Now because I might want to use the mean  and the number observations later, I'm going to do this as a very short "program." First I'm going to create a new object called *my_observations* with the value of 6. Then I'm going create an object called *my_mean* and compute the value of that using *my_sum* and *my_observations*. Then I'll return the value of the object to the screen:

```{r}
#create an object with the count of data points or observations
my_observations <- 6

#create an object called my mean then assign using <- 
#the value of my_sum divided by my_observations to my_mean
my_mean <- my_sum / my_observations

#print the value of my_mean to the screen
my_mean 

```

On your own, compute the mean of:

10,16,19,21,25,27,28

**Some of the comment lines are hints that you can either uncomment by removing the # or retype. For example, new_obs <-
is the first part of the first line of code. It's not a complete line though! Something has to go after the <- **

```{r}

#option 1 - add them up and divide by the number of observations
#If you use parentheses, you can do this in one line

#count the number of observations and assign it object new_obs
#new_obs <- 
  
#use math operation to assign the cum of the values to object #new_sum      
#new_sum <- 

#use math operation to create new object new_mean and assign it #the value of new_sum  divided by new_obs       
#new_mean <- 

#Return new_mean to the screen       
#new_mean

```

Jumping ahead just a bit, the reason for doing all that with creating objects becomes clear if we want to find the *sample* variance and standard deviation of the first problem:

```{r}

#5,6,7,12,13,14

#square the differences between the values and the mean, then
#add them up
my_diffs_squared <- (5 - my_mean)^2 + (6 - my_mean)^2 +
        (7 - my_mean)^2 + (12 - my_mean)^2 + (13 - my_mean)^2 +
        (14 - my_mean)^2

#divide the sum of squared differences by the number of #observations minus 1 (for sample variance)
my_variance <- my_diffs_squared / (my_observations - 1)

#print my_variance to the screen
my_variance

#checking against the R function
my_vec <- c(5,6,7,12,13,14)
my_check <- var(my_vec)
my_check

```
Now for the standard deviation:

```{r}

#standard deviation is square root of variance
my_sd <- sqrt(my_variance)

#print my_sd to the screen
my_sd

#checking it against the R function
my_sd_check <- sd(my_vec)
my_sd_check
```

