---
title: 'Problem Set 2: Due Wednesday July 20, 2022, 11:59 PM'
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y, %H : %M : %S')`"
output:
  html_document:
    df_print: paged
---

## Problem Set 2

### Remember to "knit" your final work. It is a good idea to export a copy to your local computer. 

*Code is not required, unless it is convenient for you to calculate answers. You can type answers directly.*

1. In R, if I want to assign the series of values 1,2,3,4,5 to a vector or to perform a function such as *mean()* on it, I would enclose it in parentheses with something in front. Would it be: _______

        - a. function(1,2,3,4,5)
        - b. makethis(1,2,3,4,5)
        - c. c(1,2,3,4,5)
        - d. wrap(1,2,3,4,5)
        
2. A __________ is a matrix of dimension 1 x 1.

3. In R, if I have a vector x = c(5,10,15,20) and a scalar z = (5), what would the following code produce:
        x * z
        
        - a. 250
        - b. 100
        - c. 5,10,15,20
        - d. 25,50,75,100
        - e. size mismatch error

4. The ___________ measures the average squared distance of the data points from the mean. The _________ is the square root of this.

**Please compute these using a calculator or math functions in R and using the population statistic method used in class (divide by n). Remember that the var() and sd() functions will give the slightly different sample statistics.**

For the following problems, use these data points:

(0.87,-6.92,-0.49,8.17,8.33,11.29,-4.85)

5. Find the sum
```{r}
0.87 -6.92 -0.49 + 8.17 + 8.33 + 11.29 -4.85

xvector <- c(0.87,-6.92,-0.49,8.17,8.33,11.29,-4.85)

sum(c(0.87,-6.92,-0.49,8.17,8.33,11.29,-4.85))
```


6. How many observations are there? 
7
7. What is the mean?
```{r}
16.4 / 7

xmean <- mean(c(0.87,-6.92,-0.49,8.17,8.33,11.29,-4.85))
```

8. What is the variance?



```{r}

xsquares <- sum((xvector-xmean)^2)
xsquares

xsquares/7

sqrt(42.51213)


```



9. What is the standard deviation?

For the following problems, use these data points:

(-2,-3,15,0,3)

10. Find the sum
11. How many observations are there?
12. What is the mean?
13. What is the median? (Note: Put them in order first!)
14. What is the variance?

Here is a sample code chunk for your use:

```{r}

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
