---
title: |
    | Problem Set 4
    | Statistical Tests
    | POLS3316: Statistics for Political Scientists
    | Due November 15, 2023 at 11:59pm
author: "Your Name"
format: 
        html:
                self-contained: true
---


## Use this work file if working in R

It is recommended to use this .qmd file to complete your work if working in R. Simply edit your name in the second line. 

## General instructions

`#` indicates that the line is a comment and does not need to be edited.

The significance level for all tests is .05 unless otherwise specified.

Type only code in the code chunks. Do not type answers as comments in the code chunks.

Type answers in the text chunks.

You can get help with the functions using the ? command in the Console (lower left)

To complete Bonus Problem 7, you will need to install a package and load a library. Instructions are given. You will not be able to get help for the z.test using the ? until the package is installed.

## Problem 1: Chi-square

Complete a Chi-square test of independence on the following data using the R function `chisq.test()`.

| | 1 | 2 | 
| --- | --- | --- |
| A | 10 | 20 |
| B | 15 | 25 |

```{r}
# Your code here


```
X-squared:

p-value:

Is the test significant? Why or why not?


## Problem 2: Chi-square

Complete a Chi-square test of independence on the following data using the R function `chisq.test()`.

| | 1 | 2 | 3 | 4 | 
| --- | --- | --- | --- | --- |
| A | 10 | 20 | 30 | 40 |
| B | 15 | 25 | 35 | 45 |
| C | 20 | 30 | 40 | 50 |

```{r}
# Your code here

```

Chi-squared:

p-value:

Is the test significant? Why or why not?


## Problem 3: t-test

Complete a paired sample t-test on the following data using the R function `t.test()`. Complete a two-tailed test.

| | Before | After |            
| --- | --- | --- |
| A | 10 | 20 |
| B | 15 | 25 |
| C | 20 | 30 |
| D | 25 | 35 |
| E | 30 | 40 |
| F | 35 | 45 |



```{r}

# Your code here


```

t-score:

p-value:

Is the test significant? Why or why not?

## Problem 5: t-test

Complete a one-sample t-test on the following data using the R function `t.test()`. Complete a two-tailed test.

| | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 | 9 | 10 |
| --- | --- | --- | --- | --- | --- | --- | --- | --- | --- | --- |
| A | 10 | 20 | 30 | 40 | 50 | 60 | 70 | 80 | 90 | 100 |

```{r}
# your code here

```

t-score:

p-value:

Is the test significant? Why or why not?

## Problem 6: Bonus

Complete a one-way ANOVA on the following data using the R function `aov()`. Complete a two-tailed test.

| | 1 | 2 | 3 | 4 |
| --- | --- | --- | --- | --- |
| A | 10 | 20 | 30 | 40 |
| B | 15 | 25 | 35 | 45 |
| C | 20 | 30 | 40 | 50 |

```{r}
# your code here

```

Is the test significant? Why or why not?


## Problem 7: Bonus

Complete a z-test on the following data using the R function `z.test()`. Complete a two-tailed test.

| | obs 1 | obs 2 | obs 3 | obs 4 | obs 5 | obs 6                
| --- | --- | --- | --- | --- | --- | --- |
| A | 10 | 20 | 30 | 40 | 50 | 60 |
| B | 15 | 25 | 35 | 45 | 55 | 65 |              


```{r}
# To run a z.test, you will need to install the BSDA package and load the library

# To do this, remove the # from the following two lines of code and run them

#install.packages("BSDA")
#library(BSDA)


# Your code here


```
z-score here:

p-value here:

Is the test significant? Why or why not?

