---
title: "Problem Set 2: POLS 3316 - Due October 6"
author: "Author name"
format: 
        html:
                self-contained: true
---

## Problem Set 2

## Instructions and Reminders

- Remember to "render" or "knit" your final work. It is a good idea to export a copy to your local computer if using Posit Cloud. If using R Studio, you will find the finished HTML or PDF file in the same folder as your working file (this file). 

- Edit the code at the top to replace Author Name with your First and Last name. That code section is called the YAML.
- Rename your file to include your own name
- *Code is not required, unless it is convenient for you to calculate answers. You can type answers directly.*
- If you do your work by hand, please show the steps: for example, if finding the mean and the sum is 25 and the number of observations is 5, you would answer something like:

         mean   = sum / number of observations
                =  25 / 5
                = 5

## Question 1

In R, if I want to assign the series of values (75,32,14,18,21) to a vector or to perform a function such as *mean()* on it, I would enclose the numbers in parentheses with something in front of the opening parenthesis. Would this look like: 

        - a. function(75,32,14,18,21)
        - b. makethis(75,32,14,18,21)
        - c. c(75,32,14,18,21)
        - d. wrap(75,32,14,18,21)
        
Answer: 

## Question Two
        
A __________ is a matrix of dimension 1 x 1. (It contains a single value and is represented by a geometric point, so may also be considered to have a dimension of zero (0)). 

Answer: 

## Question Three

In R, if I have a vector x = c(10,20,30,40) and a scalar z = (10), what would the following code produce:

        x * z
        
        - a. 1000
        - b. 250
        - c. 100,200,300,400
        - d. 20,30,40,50
        - e. size mismatch error
        
Answer:
        
## Question Four

The ___________ measures the average squared distance of the data points from the mean. The _________ is the square root of this.

First blank:

Second blank:

## Question Five through Fourteen Instructions and Data

**Please compute these using a calculator or math functions in R and using the population statistic method used in class (divide by n). You may use the sum() fucntion to find sums and you are **strongly encouraged** to create objects to represent each step in the calculation as in part I of Problem Set I. In other words, you are encouraged to:

1. Make a vector containing the data points, (You can choose all object names yourself.)
2. Create an object to hold the sum and assign the result of sum() with the vector from number 1
3. Create an object named *n* containing the number of observations
4. Create an object for the mean containing the sum object divided by *n*
5. Continue by creating objects for the steps in finding the variance
6. Create an object for the variance and perform the operation to create the value with the ojbects from step 5 and *n*
7. Find the standard deviation by taking sqrt() of the object from step 6
2. You may refer to the answers from Problem Set 1 and use that code freely

Remember that using the built-in *var()* and *sd()* functions will give slightly different sample statistics. If you use the R functions, you will need to convert them to population statistics or your answer will be wrong.

For problems 5 through 9, use these data points: 

(0.95,-8.31,-0.41,7.22,8.33,13.79,-2.88)

For problems 10 through 14 use these data points: 

(-10,-5,18,1,5)

## Question 5

5. Find the sum

```{r}

```

Answer:

## Question 6

How many observations are there? 

Answer:

It would be useful to create an object named n in this code box and store the answer to this problem in that object:

```{r}



```

## Question 7

What is the mean?

```{r}



```

Answer:

## Question 8

What is the variance?

```{r}


```

Answer:

## Question 9

What is the standard deviation?

```{r}

```


Answer:


## Data Questions 10 to 14

For problems 10 through 14 use these data points:

(-10,-5,18,1,5)

## Question 10

Find the sum

```{r}

```


Answer:

## Question 11

How many observations are there?

Answer:

```{r}

```

##Question 12

What is the mean?

```{r}

```

Answer: 

## Question 13

What is the median? (Note: If you do not use R code, remember to put them in order first! )

```{r}

```

Answer:


## Question 14

What is the variance?

```{r}

```

Answer: 


## Question 15 (Bonus)

What is the standard deviation?

```{r}

```

Answer: 

